\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{caption}

\geometry{margin=2.5cm}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{red},
    frame=single,
    breaklines=true,
    numbers=left,
    numberstyle=\tiny\color{gray},
    backgroundcolor=\color{gray!5},
}

\title{Calculating Peak GPU Memory Usage from Code Constants:\\
       GPT-2 with GPipe Pipeline Parallelism}
\author{}
\date{}

\begin{document}
\maketitle

\section{Introduction}

This document describes how to calculate the theoretical peak GPU memory usage
for the GPT-2 pipeline-parallel training script
\texttt{gpt2\_gpipe\_memory\_profile.py} using only the constants defined in
the source code. We derive closed-form formulas for each memory component and
show how they compose to predict the observed peak allocation of
$\approx 625.77$\,MB on GPU\,0 (from the example trace
\texttt{peak\_alloc\_events\_GPU0\_for example.json}, where
$\texttt{max\_allocated\_mem} = 656\,165\,888$ bytes).

%===========================================================================
\section{Code Constants}
\label{sec:constants}

All quantities below are drawn directly from the source file.

\subsection{Training Constants}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Symbol} & \textbf{Code Variable} & \textbf{Value} \\
\midrule
$B$             & \texttt{BATCH\_SIZE}       & 8     \\
$G$             & \texttt{NUM\_GPUS}         & \texttt{torch.cuda.device\_count()} \\
$C$             & \texttt{MICRO\_BATCHES}    & $G$   \\
$S$             & \texttt{SEQ\_LEN}          & 256   \\
$T$             & \texttt{NUM\_STEPS}        & 10    \\
\bottomrule
\end{tabular}
\end{center}

\subsection{GPT-2 Model Constants}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Symbol} & \textbf{Code Variable / Config Key} & \textbf{Value} \\
\midrule
$V$       & \texttt{vocab\_size}              & 50\,257  \\
$P$       & \texttt{n\_positions}             & 1\,024   \\
$d$       & \texttt{n\_embd}                  & 768      \\
$L$       & \texttt{n\_layer}                 & 12       \\
$H$       & \texttt{n\_head}                  & 12       \\
$d_h$     & $d / H$                           & 64       \\
$d_{\text{ff}}$ & \texttt{n\_inner}           & 3\,072   \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Derived Constants}

\begin{itemize}
    \item \textbf{Bytes per parameter}: $b = 4$ (all parameters are \texttt{float32}).
    \item \textbf{Micro-batch size}: $B_\mu = B / C = B / G$.
    \item \textbf{Total sequential modules}: $M = 1 + L + 1 = 14$
          (1 \texttt{EmbeddingBlock} $+$ 12 \texttt{TransformerBlock} $+$ 1 \texttt{LMHead}).
    \item \textbf{Balance per GPU}: the code distributes $M$ modules by
          \[
              \text{balance}[i] =
              \left\lfloor \frac{M}{G} \right\rfloor
              + \mathbf{1}\!\left[i < M \bmod G\right],
              \quad i = 0, \dots, G-1.
          \]
\end{itemize}

%===========================================================================
\section{Memory Components}
\label{sec:components}

GPU memory during training is consumed by five main categories:

\begin{equation}
\label{eq:total}
\boxed{
    \mathcal{M}_{\text{peak}}^{(i)}
    \;=\;
    \mathcal{M}_{\text{param}}^{(i)}
    + \mathcal{M}_{\text{grad}}^{(i)}
    + \mathcal{M}_{\text{opt}}^{(i)}
    + \mathcal{M}_{\text{act}}^{(i)}
    + \mathcal{M}_{\text{misc}}^{(i)}
}
\end{equation}

where $i$ denotes the GPU index.

%-------------------------------------------------------------------
\subsection{Model Parameters \texorpdfstring{$\mathcal{M}_{\text{param}}$}{M\_param}}
\label{sec:params}

\subsubsection{EmbeddingBlock (Module 0)}

\begin{align}
    \theta_{\texttt{wte}} &= V \times d = 50\,257 \times 768 = 38\,597\,376 \\
    \theta_{\texttt{wpe}} &= P \times d = 1\,024 \times 768 = 786\,432 \\
    \theta_{\text{emb}}   &= \theta_{\texttt{wte}} + \theta_{\texttt{wpe}} = 39\,383\,808
\end{align}

\subsubsection{Single TransformerBlock (Modules 1--12)}

Each transformer block contains:

\begin{center}
\begin{tabular}{llr}
\toprule
\textbf{Sub-module} & \textbf{Formula} & \textbf{Parameters} \\
\midrule
\texttt{ln\_1} (LayerNorm)
    & $2d$ (weight $+$ bias)
    & 1\,536 \\
\texttt{c\_attn} (Linear)
    & $d \times 3d + 3d$
    & 1\,771\,776 \\
\texttt{c\_proj} (Linear, Attn)
    & $d \times d + d$
    & 590\,592 \\
\texttt{ln\_2} (LayerNorm)
    & $2d$
    & 1\,536 \\
\texttt{c\_fc} (Linear, MLP)
    & $d \times d_{\text{ff}} + d_{\text{ff}}$
    & 2\,362\,368 \\
\texttt{c\_proj} (Linear, MLP)
    & $d_{\text{ff}} \times d + d$
    & 2\,360\,064 \\
\midrule
\textbf{Total per block} $\theta_{\text{block}}$
    & $2(2d) + d(3d) + 3d + d^2 + d + d \cdot d_{\text{ff}} + d_{\text{ff}} + d_{\text{ff}} \cdot d + d$
    & \textbf{7\,087\,872} \\
\bottomrule
\end{tabular}
\end{center}

In closed form:
\begin{equation}
    \theta_{\text{block}} = 4d + 3d^2 + 3d + d^2 + d + d \cdot d_{\text{ff}} + d_{\text{ff}} + d_{\text{ff}} \cdot d + d
    = 4d^2 + 2d \cdot d_{\text{ff}} + d_{\text{ff}} + 9d
\end{equation}

\subsubsection{LMHead (Module 13)}

\begin{align}
    \theta_{\texttt{ln\_f}} &= 2d = 1\,536 \\
    \theta_{\texttt{lm\_head}} &= d \times V = 768 \times 50\,257 = 38\,597\,376 \quad \text{(no bias)} \\
    \theta_{\text{head}} &= 2d + d \cdot V = 38\,598\,912
\end{align}

\subsubsection{Total Model Parameters}

\begin{equation}
\boxed{
    \Theta = \theta_{\text{emb}} + L \cdot \theta_{\text{block}} + \theta_{\text{head}}
    = 39\,383\,808 + 12 \times 7\,087\,872 + 38\,598\,912
    = 163\,037\,184 \;\approx\; 163\text{M}
}
\end{equation}

\begin{equation}
    \mathcal{M}_{\text{param}}^{\text{total}} = \Theta \times b = 163\,037\,184 \times 4 = 652\,148\,736 \text{ bytes} \approx 621.94 \text{ MB}
\end{equation}

%-------------------------------------------------------------------
\subsection{Per-GPU Parameter Memory (Pipeline Split)}
\label{sec:pergpu}

GPipe distributes the $M = 14$ sequential modules across $G$ GPUs according to
the balance vector. Let $L_i$ denote the number of \emph{transformer blocks}
assigned to GPU $i$, with the EmbeddingBlock always on GPU\,0 and LMHead on
GPU\,$G{-}1$.

\begin{equation}
    \Theta^{(i)} = \mathbf{1}[i=0]\;\theta_{\text{emb}}
                 + L_i \cdot \theta_{\text{block}}
                 + \mathbf{1}[i=G{-}1]\;\theta_{\text{head}}
\end{equation}

\begin{equation}
    \mathcal{M}_{\text{param}}^{(i)} = \Theta^{(i)} \times b
\end{equation}

\paragraph{Example with $G = 2$:}
Balance $= [7, 7]$. GPU\,0 holds modules 0--6 (EmbeddingBlock $+$ 6 TransformerBlocks);
GPU\,1 holds modules 7--13 (6 TransformerBlocks $+$ LMHead).

\begin{align}
    \Theta^{(0)} &= 39\,383\,808 + 6 \times 7\,087\,872 = 81\,911\,040
    &\Rightarrow \mathcal{M}_{\text{param}}^{(0)} &= 312.47\text{ MB} \\
    \Theta^{(1)} &= 6 \times 7\,087\,872 + 38\,598\,912 = 81\,126\,144
    &\Rightarrow \mathcal{M}_{\text{param}}^{(1)} &= 309.47\text{ MB}
\end{align}

\paragraph{Example with $G = 4$:}
Balance $= [4, 3, 3, 4]$. GPU\,0 holds EmbeddingBlock~$+$~3 TransformerBlocks;
GPU\,3 holds 3 TransformerBlocks~$+$~LMHead.

\begin{align}
    \Theta^{(0)} &= 39\,383\,808 + 3 \times 7\,087\,872 = 60\,647\,424
    &\Rightarrow \mathcal{M}_{\text{param}}^{(0)} &= 231.34\text{ MB}
\end{align}

%-------------------------------------------------------------------
\subsection{Gradient Memory \texorpdfstring{$\mathcal{M}_{\text{grad}}$}{M\_grad}}

During the backward pass every parameter accumulates a gradient of the
same dtype (\texttt{float32}):

\begin{equation}
    \mathcal{M}_{\text{grad}}^{(i)} = \Theta^{(i)} \times b
    = \mathcal{M}_{\text{param}}^{(i)}
\end{equation}

Gradients are allocated lazily during the first backward pass and persist
until \texttt{optimizer.zero\_grad()} is called.

%-------------------------------------------------------------------
\subsection{Optimizer State Memory \texorpdfstring{$\mathcal{M}_{\text{opt}}$}{M\_opt}}

AdamW maintains two auxiliary states per parameter (first moment $m$ and
second moment $v$), each of the same size as the parameter:

\begin{equation}
    \mathcal{M}_{\text{opt}}^{(i)} = 2 \times \Theta^{(i)} \times b
    = 2 \times \mathcal{M}_{\text{param}}^{(i)}
\end{equation}

These states are allocated \textbf{lazily on the first call} to
\texttt{optimizer.step()}, i.e., after the first backward pass completes.
Before the first step, $\mathcal{M}_{\text{opt}}^{(i)} = 0$.

%-------------------------------------------------------------------
\subsection{Activation Memory \texorpdfstring{$\mathcal{M}_{\text{act}}$}{M\_act}}
\label{sec:activations}

Activation memory is affected by two key features of GPipe:

\begin{enumerate}
    \item \textbf{Micro-batching}: the batch of size $B$ is split into
          $C = G$ chunks, each of effective size $B_\mu = B / G$.
    \item \textbf{Activation checkpointing} (\texttt{checkpoint='always'}):
          intermediate activations within each micro-batch partition are
          \emph{not} saved; instead, only the input tensor to each
          checkpointed segment is retained, and intermediates are
          recomputed during the backward pass.
\end{enumerate}

\subsubsection{Without Checkpointing (Reference)}

For a single transformer block processing one micro-batch, the peak
intermediate activations include:

\begin{center}
\begin{tabular}{llr}
\toprule
\textbf{Tensor} & \textbf{Shape} & \textbf{Size (bytes)} \\
\midrule
Hidden states (input)
    & $B_\mu \times S \times d$
    & $B_\mu \cdot S \cdot d \cdot b$ \\
QKV projection
    & $B_\mu \times S \times 3d$
    & $3 \, B_\mu \cdot S \cdot d \cdot b$ \\
Attention scores
    & $B_\mu \times H \times S \times S$
    & $B_\mu \cdot H \cdot S^2 \cdot b$ \\
Attention output
    & $B_\mu \times S \times d$
    & $B_\mu \cdot S \cdot d \cdot b$ \\
MLP intermediate
    & $B_\mu \times S \times d_{\text{ff}}$
    & $B_\mu \cdot S \cdot d_{\text{ff}} \cdot b$ \\
\bottomrule
\end{tabular}
\end{center}

Per layer, per micro-batch (without checkpointing):
\begin{equation}
    \mathcal{A}_{\text{layer}} =
    B_\mu \cdot S \cdot b \left(
        d + 3d + d + d_{\text{ff}}
    \right)
    + B_\mu \cdot H \cdot S^2 \cdot b
\end{equation}

With $B_\mu = 4$, $S = 256$, $d = 768$, $d_{\text{ff}} = 3072$, $H = 12$:
\begin{equation}
    \mathcal{A}_{\text{layer}} = 4 \times 256 \times 4 \times (768 + 2304 + 768 + 3072)
    + 4 \times 12 \times 256^2 \times 4
    \approx 31.50\text{ MB}
\end{equation}

\subsubsection{With Checkpointing (as Used)}

With \texttt{checkpoint='always'}, GPipe only stores the \textbf{input tensor}
to each micro-batch partition during the forward pass. All internal
activations are recomputed during the backward pass for one segment at a time.
The stored activation per layer per micro-batch reduces to:

\begin{equation}
    \mathcal{A}_{\text{ckpt}} = B_\mu \cdot S \cdot d \cdot b
\end{equation}

With $B_\mu = 4$, $S = 256$, $d = 768$, $b = 4$:
\begin{equation}
    \mathcal{A}_{\text{ckpt}} = 4 \times 256 \times 768 \times 4 = 3\,145\,728 \text{ bytes} = 3.0\text{ MB per layer}
\end{equation}

During the backward pass, one segment's activations are recomputed at a
time, and per-layer peak is:
\begin{equation}
    \mathcal{A}_{\text{recompute}} = \mathcal{A}_{\text{layer}} \approx 31.50\text{ MB}
    \quad \text{(transient, for one layer at a time)}
\end{equation}

The total activation memory on GPU $i$ during steady-state pipeline execution
with $C$ micro-batches is bounded by:
\begin{equation}
    \mathcal{M}_{\text{act}}^{(i)} \leq
    C \cdot L_i \cdot \mathcal{A}_{\text{ckpt}}
    + \mathcal{A}_{\text{recompute}}
\end{equation}

The first term accounts for the saved checkpoint inputs across all
micro-batches and layers on that GPU; the second term is the transient
recomputation cost for the currently active backward segment.

%-------------------------------------------------------------------
\subsection{Miscellaneous Memory \texorpdfstring{$\mathcal{M}_{\text{misc}}$}{M\_misc}}

Additional allocations include:

\begin{itemize}
    \item \textbf{Input tensor}: $B \times S \times 8$ bytes
          (dtype \texttt{int64}) $= 8 \times 256 \times 8 = 16\,384$ bytes $\approx 0.016$\,MB.
    \item \textbf{Labels tensor}: same size as input, on last GPU.
    \item \textbf{Logits tensor}: $B \times S \times V \times b$ on last GPU
          $= 8 \times 256 \times 50\,257 \times 4 \approx 391.85$\,MB.
    \item \textbf{Causal mask}: $S \times S \times 1$ byte (\texttt{bool})
          $= 256^2 = 65\,536$ bytes, per layer evaluation.
    \item \textbf{CUDA context and allocator overhead}: typically 200--500\,MB
          (loaded once per GPU, not reflected in \texttt{memory\_allocated}).
    \item \textbf{Profiler buffers}: memory history recording
          (\texttt{\_record\_memory\_history}) uses additional host and device memory.
\end{itemize}

%===========================================================================
\section{Peak Memory Formula}
\label{sec:peak}

The peak memory on GPU $i$ occurs at different training phases.  Below we
analyze the two critical moments.

\subsection{Phase 1: After Model Load, Before Training}

Only parameters are allocated:
\begin{equation}
    \mathcal{M}_{\text{phase1}}^{(i)} = \mathcal{M}_{\text{param}}^{(i)}
\end{equation}

\subsection{Phase 2: During First Backward Pass (Before First \texttt{optimizer.step()})}

At this point, parameters and gradients are live, plus transient
activations from checkpointing recomputation:
\begin{equation}
    \mathcal{M}_{\text{phase2}}^{(i)}
    = \mathcal{M}_{\text{param}}^{(i)}
    + \mathcal{M}_{\text{grad}}^{(i)}
    + \mathcal{M}_{\text{act}}^{(i)}
\end{equation}

\subsection{Phase 3: Steady State (After First \texttt{optimizer.step()})}

The optimizer states now exist, but gradients may be zeroed:
\begin{equation}
    \mathcal{M}_{\text{phase3}}^{(i)}
    = \mathcal{M}_{\text{param}}^{(i)}
    + \mathcal{M}_{\text{grad}}^{(i)}
    + \mathcal{M}_{\text{opt}}^{(i)}
    + \mathcal{M}_{\text{act}}^{(i)}
\end{equation}

\subsection{Overall Peak}

The overall peak is achieved during steady-state backward passes (Phase~3):
\begin{equation}
\boxed{
    \mathcal{M}_{\text{peak}}^{(i)}
    = 4 \, \Theta^{(i)} \cdot b
    + C \cdot L_i \cdot B_\mu \cdot S \cdot d \cdot b
    + \mathcal{A}_{\text{recompute}}
    + \mathcal{M}_{\text{misc}}^{(i)}
}
\end{equation}

where the factor of 4 comes from: parameters ($1\times$) + gradients ($1\times$) +
optimizer first moment ($1\times$) + optimizer second moment ($1\times$).

%===========================================================================
\section{Numerical Example}
\label{sec:numerical}

We compute the peak for \textbf{GPU\,0} using the code's default constants
and illustrate with $G = 2$ (the minimum required).

\subsection{Step-by-Step Calculation}

\begin{enumerate}
    \item \textbf{Balance}: $M = 14$, $G = 2 \Rightarrow$ balance $= [7, 7]$.
          GPU\,0 holds modules 0--6: EmbeddingBlock $+$ 6 TransformerBlocks, so $L_0 = 6$.

    \item \textbf{Parameters on GPU\,0}:
          \begin{equation}
              \Theta^{(0)} = \underbrace{39\,383\,808}_{\theta_{\text{emb}}}
              + 6 \times \underbrace{7\,087\,872}_{\theta_{\text{block}}}
              = 81\,911\,040 \text{ parameters}
          \end{equation}

    \item \textbf{Parameter memory}:
          $\mathcal{M}_{\text{param}}^{(0)} = 81\,911\,040 \times 4 = 327\,644\,160$ bytes $= 312.47$ MB.

    \item \textbf{Gradient memory}:
          $\mathcal{M}_{\text{grad}}^{(0)} = 312.47$ MB.

    \item \textbf{Optimizer state memory}:
          $\mathcal{M}_{\text{opt}}^{(0)} = 2 \times 312.47 = 624.93$ MB.

    \item \textbf{Checkpoint activation memory} ($C = 2$ micro-batches, $L_0 = 6$ layers, $B_\mu = 4$):
          \begin{equation}
              C \cdot L_0 \cdot \mathcal{A}_{\text{ckpt}}
              = 2 \times 6 \times (4 \times 256 \times 768 \times 4)
              = 2 \times 6 \times 3\,145\,728 = 37\,748\,736 \text{ bytes}
              = 36.0\text{ MB}
          \end{equation}

    \item \textbf{Recomputation transient}: $\mathcal{A}_{\text{recompute}} \approx 31.50$ MB (one layer at a time).
\end{enumerate}

\subsection{Summary Table for GPU\,0 ($G = 2$)}

\begin{center}
\begin{tabular}{lr}
\toprule
\textbf{Component} & \textbf{Size (MB)} \\
\midrule
Parameters $\mathcal{M}_{\text{param}}^{(0)}$               & 312.47 \\
Gradients $\mathcal{M}_{\text{grad}}^{(0)}$                 & 312.47 \\
Optimizer states $\mathcal{M}_{\text{opt}}^{(0)}$            & 624.93 \\
Checkpoint saved activations                                  & 36.00  \\
Recompute transient activations                               & 31.50  \\
Input tensor + causal masks                                   & $\sim$0.08 \\
\midrule
\textbf{Estimated peak}                                       & $\sim$\textbf{1\,317.45} \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textit{Note}: The steady-state peak includes all four copies
(params $+$ grads $+$ $m$ $+$ $v$).  In practice, the observed peak may be
lower if \texttt{optimizer.zero\_grad()} has freed gradient memory before the
optimizer states coexist with new gradients, or if the optimizer states have
not yet been allocated (first forward/backward only).

%===========================================================================
\section{Observed Peak and Interpretation}
\label{sec:observed}

The example trace \texttt{peak\_alloc\_events\_GPU0\_for example.json}
records a peak of:
\[
    \texttt{max\_allocated\_mem} = 656\,165\,888 \text{ bytes} \approx 625.77\text{ MB}
\]

This value closely matches $2\times \mathcal{M}_{\text{param}}^{(0)} \approx 624.93$\,MB,
suggesting that \textbf{the peak occurs during the backward pass when both
parameters and gradients are live}, but \textbf{before the first
\texttt{optimizer.step()}} allocates the Adam states (Phase~2).  The small
excess ($\sim$0.84\,MB) corresponds to transient activation tensors and
the input data.

This can be expressed as:
\begin{equation}
\boxed{
    \mathcal{M}_{\text{observed peak}}^{(0)} \approx 2\,\Theta^{(0)} \cdot b + \epsilon
    = 2 \times 81\,911\,040 \times 4 + \epsilon
    \approx 625.77\text{ MB}
}
\end{equation}

where $\epsilon$ captures transient activations ($\sim$0.84\,MB in this trace).

After the first \texttt{optimizer.step()}, the Adam states add another
$2\times\mathcal{M}_{\text{param}}^{(0)} \approx 625$\,MB, but subsequent
peak measurements via \texttt{max\_memory\_allocated} would reflect that
higher plateau.

%===========================================================================
\section{General Formulas (Quick Reference)}
\label{sec:summary}

Let GPU $i$ hold an embedding indicator $e_i \in \{0,1\}$, a head indicator
$h_i \in \{0,1\}$, and $L_i$ transformer blocks.  Then:

\begin{align}
    \Theta^{(i)}
    &= e_i (Vd + Pd) + L_i (4d^2 + 2d\,d_{\text{ff}} + d_{\text{ff}} + 9d)
       + h_i (2d + Vd) \\[6pt]
    \mathcal{M}_{\text{param}}^{(i)} &= \Theta^{(i)} \cdot b \\[3pt]
    \mathcal{M}_{\text{grad}}^{(i)}  &= \Theta^{(i)} \cdot b \\[3pt]
    \mathcal{M}_{\text{opt}}^{(i)}   &= 2\,\Theta^{(i)} \cdot b
                                        \quad\text{(AdamW: $m + v$)} \\[3pt]
    \mathcal{M}_{\text{act}}^{(i)}   &\leq C \cdot L_i \cdot B_\mu \cdot S \cdot d \cdot b
                                        + B_\mu \cdot S \cdot b\,(5d + d_{\text{ff}})
                                        + B_\mu \cdot H \cdot S^2 \cdot b
\end{align}

\begin{equation}
\boxed{
    \mathcal{M}_{\text{peak}}^{(i)}
    = \underbrace{4\,\Theta^{(i)} b}_{\text{params + grads + optimizer}}
    + \underbrace{C \cdot L_i \cdot B_\mu S d\, b}_{\text{ckpt activations}}
    + \underbrace{B_\mu S\, b\,(5d + d_{\text{ff}}) + B_\mu H S^2 b}_{\text{recompute transient}}
    + \mathcal{M}_{\text{misc}}^{(i)}
}
\end{equation}

\subsection{Substitution Table}

To evaluate for a specific configuration, substitute:

\begin{center}
\begin{tabular}{l|c|c}
\toprule
\textbf{Constant} & \textbf{Code location} & \textbf{Default} \\
\midrule
$B$             & Line 30: \texttt{BATCH\_SIZE = 8}                  & 8 \\
$G$             & Line 29: \texttt{NUM\_GPUS = torch.cuda.device\_count()} & runtime \\
$S$             & Line 32: \texttt{SEQ\_LEN = 256}                   & 256 \\
$V$             & Line 131: \texttt{vocab\_size=50257}               & 50\,257 \\
$P$             & Line 132: \texttt{n\_positions=1024}               & 1\,024 \\
$d$             & Line 133: \texttt{n\_embd=768}                     & 768 \\
$L$             & Line 134: \texttt{n\_layer=12}                     & 12 \\
$H$             & Line 135: \texttt{n\_head=12}                      & 12 \\
$d_{\text{ff}}$ & Line 136: \texttt{n\_inner=3072}                   & 3\,072 \\
$b$             & (implicit \texttt{float32})                        & 4 \\
\bottomrule
\end{tabular}
\end{center}

%===========================================================================
\section{Conclusion}

By expressing every memory component in terms of the code constants
$B$, $G$, $S$, $V$, $P$, $d$, $L$, $H$, $d_{\text{ff}}$, and $b$,
we can predict the peak GPU memory usage \emph{before running the script}.
The dominant terms are the model parameters and their associated gradients
and optimizer states ($4\Theta^{(i)} b$), which scale linearly with model
size and inversely with the number of GPUs.  Activation memory is
significantly reduced by GPipe's built-in checkpointing
(\texttt{checkpoint='always'}), leaving only $\mathcal{O}(B_\mu \cdot S \cdot d)$
per saved checkpoint.  The observed peak of $\approx 625.77$\,MB on GPU\,0
is consistent with $2\Theta^{(0)} b$ (parameters plus gradients during the
first backward pass, before optimizer state allocation).

\end{document}
